Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-07-01 Experiments.tex
Down to 10 pages

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
        \vspace{-0.3in}
    \caption{Strong scalability:  average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
\label{fig:strongScala}
\end{figure}

>>>>

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
        \vspace{-0.3in}
    \caption{Strong scalability: Average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
\label{fig:strongScala}
\end{figure}

### 2017-07-01 nge.tex
Down to 10 pages

As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
as Titan's worker nodes lack connectivity to the wide area network The lack of
pilot capabilities impacts both the efficiency and the flexibility of PanDA's
execution process. Pilots could improve efficiency by increasing throughput and
enabling greater backfill utilization. Further, pilots makes it easier to
support heterogeneous workloads.

The absence of pilots imposes the static coupling between MPI scripts
submitted to the PBS batch system and detector simulations
(\S\ref{sec:panda_titan}). This makes the scheduling of multiple generations
of workload on the same PBS job impossible: once a statically defined number
of detector simulations are packaged into a PBS job and this job is queued on
Titan, no further simulations can be added to that job. New simulations have
to be packaged into a new PBS job that needs to be submitted to Titan based
upon backfill availability.

The support of  multiple generations of workload would enable more efficient use
of the backfill availability of walltime. Currently, when a set of simulations
ends, the PBS job also ends, independent of whether more wall-time would still
be available. With a pilot, additional simulations could be executed  to utilize
all the available wall-time, while avoiding further job packaging and submission
overheads.

@@

% -----------------------------------------------------------------------------
\subsection{Implementation}
\label{sec:arch}

The implementation of pilot capabilities within the current PanDA Broker require
quantification of the effective benefits that it could yield and, on the base of
this analysis, a dedicated engineering effort. We developed a prototype of a
pilot system capable of executing on Titan to study experimentally the
quantitative and qualitative benefits that it could bring to PanDA. We called
this prototype Next Generation Executor (NGE).

@@

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
   \vspace{-0.3in}
  \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan
  DTN while the Agent is executed on its worker nodes. Color coding: gray for
  entities external to NGE, white for APIs, purple for NGE's modules, green for
  pilots, yellow for module's components.}
\label{fig:arch-overview}
\end{figure}

NGE exposes an API to describe workloads (Fig.~\ref{fig:arch-overview}, green
squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
instantiate a PilotManager and a UnitManager.   The PilotManager submits
pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
the Agent's Executors on the on the worker
nodes, and the UnitManager schedules units to the Agent's Scheduler
(Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's Scheduler schedules
the units on one or more Agent's Executor for execution. The Agent's executors
can manage one or more worker nodes, depending on performance evaluations. The
UnitManager and the Agent communicate via a database that is instantiated on
one of Titan's DTN so as to be reachable by both modules.

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
communication and coordination of the execution of units. This environment is
a critical component of the OpenMPI implementation, developed to support
distributed high-performance computing applications operating in a
heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
multiple nodes. The DVM transparently provides support for interprocess
communication, resource discovery and allocation, and process launching across
a variety of platforms. Libraries enable the interaction between users and
ORTE to enable the submission, monitoring and managing of tasks, avoiding
filesystem bottlenecks and race conditions with network sockets. As a
consequence, ORTE is able to minimize the system overhead while submitting
tasks.

>>>>

% As seen in \S\ref{ssec:panda_titan}, PanDA Broker is deployed on the DTNs
% as Titan's worker nodes lack connectivity to the wide area network.
The lack of pilot capabilities in PanDA Broker impacts both the efficiency
and the flexibility of PanDA's execution process. Pilots could improve
efficiency by increasing throughput and enabling greater backfill
utilization. Further, pilots would make it easier to support heterogeneous
workloads.

The absence of pilots 
% imposes the static coupling between MPI scripts submitted to the PBS batch
% system and detector simulations (\S\ref{sec:panda_titan}). This
makes the scheduling of multiple generations of workload on the same PBS job
impossible: once a statically defined number of detector simulations are
packaged into a PBS job and this job is queued on Titan, no further
simulations can be added to that job. New simulations have to be packaged
into a new PBS job that needs to be submitted to Titan based upon backfill
and PanDA Brokers availability.

The support of multiple generations of workload would enable more efficient
use of the backfill availability walltime. Currently, when a set of
simulations ends, the PBS job also ends, independent of whether more
wall-time would still be available. With a pilot, additional simulations
could be executed to utilize all the available wall-time, while avoiding
further job packaging and submission overheads.

@@

% ---------------------------------------------------------------------------
\subsection{Implementation}
\label{sec:arch}

The implementation of pilot capabilities within the PanDA Broker require
quantification of the effective benefits that it could yield and, on the base
of this analysis, a dedicated engineering effort. We developed a prototype of
a pilot system capable of executing on Titan to study experimentally the
quantitative and qualitative benefits that it could bring to PanDA\@. We
called this prototype Next Generation Executor (NGE).

@@

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
   \vspace{-0.3in}
  \caption{NGE Architecture: The PilotManager and UnitManager reside on a DTN
  while the Agent is executed on a worker node. Color coding: gray for
  entities external to NGE\@; white for APIs; purple for NGE's modules; green
  for pilots; yellow for module's components.}
\label{fig:arch-overview}
\end{figure}

NGE exposes an API to describe workloads (Fig.~\ref{fig:arch-overview}, green
squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
instantiate a PilotManager and a UnitManager. The PilotManager submits pilots
to Titan's PBS batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
arrow). %, as done by PanDA Broker to submit MPI scripts. 
Once scheduled, the pilot's Agent is bootstrapped on a MOM node and the
Agent's Executors on the worker nodes.

The UnitManager and the Agent communicate via a database instantiated on a
DTN so as to be reachable by both modules. The UnitManager schedules units to
the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow) and the
Agent's Scheduler schedules the units on one or more Agent's Executor.
% for execution. The Agent's executors can manage one or more worker nodes,
% depending on performance evaluations.

The NGE Agent uses the Open Run-Time Environment (ORTE) for communication and
coordination of the execution of units. This environment is a critical
component of the OpenMPI implementation~\cite{castain05:_open_rte}.
%, cug-2016}.
ORTE is able to minimize the system overhead while submitting tasks by
avoiding filesystem bottlenecks and race conditions with network sockets.
